### 2025-04-23 docs/Homeownership_BankOfMomAndDad.tex
First pass at new empirical section (too long!)

To quantify the impact of parental transfers on homeownership, I then combine two distinct models: housing models, which ignore parental transfers, and models of parental transfers, which ignore housing. I build a life-cycle overlapping generations model with altruistic parents and a rent-or-own decision. Parents and children interact without commitment through both inter-vivos transfers and end-of-life bequests. In models of altruism without housing (e.g., \citealp{Altonji1997a,Barczyk2014}), the transfer motive is to increase the child's consumption when they are borrowing constrained, thus having high marginal utility of wealth. The inclusion of a frictional housing market generates two new transfer motives. First, transfers can directly alleviate credit constraints, enabling wealthier parents to help their children meet down-payment requirements. Second, future transfers provide partial insurance against future income shocks, reducing risks associated with the large illiquid investment of homeownership. Moreover, the illiquidity of housing serves as a commitment device enabling children to secure future transfers. As in all transfer models, the expectation of future transfers reduces children's saving motives, potentially lowering the homeownership rate.

@@

	 {\begin{footnotesize} \textit{Notes:} The left panel uses data from the SHED with the sample restricted to current owners who report this being their first home. The right panel uses data from the AHS with the sample restricted to all owner-occupied units that did not indicate the sale of a previous home as the main source of funding. Note that the left figure has year of purchase on the horizontal axis while the right figure has survey year. Dashed lines denote 95\% confidence intervals.\end{footnotesize}}
\end{figure}

\textit{SHED:} I use SHED, an annual cross-sectional survey conducted by the Federal Reserve, to observe the share of households who funded the down payment with a loan or gift from family or friends, by year of purchase. The results are plotted in the left panel of Figure \ref{fig:motivation}. The main observation is the large increase in the role of inter-vivos transfers for homeowners since 2001. From 2001 to 2007, only 10\% to 18\% of first-time buyers received transfers, while 20\% to 40\% received transfers after 2009.\footnote{The question was only included in the 2015 and 2016 waves. In the 2015 wave, first-time ownership and down-payment source is only reported for those who bought in 2001 to 2015, while in the 2016 wave it is only reported for 2015 to 2016. }

\textit{AHS:} I use the AHS, which surveys occupants of owner-occupied units, to obtain a time series on whether the down payment was mostly funded by gifts or inheritances (right panel of Figure \ref{fig:motivation}). We observe a flat pattern in the 1990s, followed by a sharp decline from 2005 to 2009, coinciding with the prevalence of mortgages featuring very low down payments (e.g., \cite{Corbae2015}), and a stark increase from 2013 to today. Minor fluctuations in the percentage of homeowners relying on gifts or inheritances reflect significant shifts among new homeowners, who comprise a small fraction of all homeowners. This is because the horizontal axis denotes the survey year, not the year of purchase.

@@

\textcolor{red}{In all other waves, there is a general question of gifts and inhertitances from people outside the hosuehold. We do not know the source of these transfers (e.g., they could be from friends or an inheritance from grandparents) and hence these do not know the extent to which these captures parental assistance. Yet, they, to some extent, capture the impact of transfer receipt on household decisions, and hence I use these variables to capture the role of transfers on housing outcomes, following \cite{Lee2018}.}

@@

I now utilize microdata from the PSID to examine the empirical association between transfer receipt and parental financial resources on housing outcomes.

I use the sum of parental net worth and income (denoted as wealth) as a statistic to measure parents' financial resources in all specifications for two reasons. First, this is consistent with models of altruism, as the one I use later. Second, since parents' wealth and income are highly correlated and I have few observations---under 1,000 for most specifications---the standard errors are large when both are included. The sample is limited to households aged 25 to 44.

\subsubsection{Transfers and the Rent-to-Own Transition}
\textcolor{red}{!! [[rewrite this]] !!}
As discussed in the introduction, several papers look at the role of parental transfers and resources for entering into the housing market \cite[see e.g.,][]{wold2024housing,Lee2019,Blickle2019}. The goal of this section is not to `re-invent' the wheel, but to simply show that these relationships look like in my data. I run regressions on renting households aged 25-44 who have not been observed as homeowners before. In the first set of regressions, the sample are all households aged 25-44 in 2011 who are renting with non-missing parental transfer data from the 2013 Transfer supplement. The outcome variable is whether the household is an owner in the next period, and the set of control variables are (log) parental cash-on-hand, (log) household wealth, (log) household income, household marital states, (log) household size, education, age, parental age, and fixed effects for race, year, and state. The results of the transfer coefficients are reported in Table \ref{tab:newowners_short} while the full regression output is in Appendix Table \ref{tab:newowners}

Column 1 reports the effect of recieving a parental transfers (above \$500) (in year 2012) on the probability of becoming a homeowner in 2013, which increases by 1.6 pp. Next, I also include the inheritance/transfer dummy, (hence includes households who received transfers/inheritances above \$10,000 in 2011, 2012, or 2013). Column 2 also includes households who report having recieved transfers or inheritances above \$10,000 while column (3) only denotes those who have received transfers above \$10k (whether transfers from parents or any inheritance). Renters who receive transfers are much more likely to transition to homeowners, even after controlling for other observable characteristics. If we only include large transfers (above \$10,000) transfer receipt is associated with a 7.7 percentage point increase in the probability of transitioning into homeownership, more than half of the transition rate. The coefficents are not statistically significant, perhaps not surprising given the small sample size, the few households who receive transfers and the few households who become homeowners.

To increase statistical power I run the same regression using the same specification, but over the entire sample period (1999-2021) using OLS and household fixed effects (FE). Using fixed effects also allows us to control for household fixed effects such as genetics, risk aversion, etc. The downside is that we only observe larger transfers (above 10k) and that we do not know the source of the transfer. If we focus on the FE estimator, we see that transfer receipt is associated with an 8.3 percentage point increase (statistically signifacant at the 10\% level) in the probability of becoming a homeowner, relative to a benchmark probability of 18.1\%.

\textcolor{red}{Missing regression equation}
It is perhaps not so surprising that the effect of transfer receipt is so big. First, beyond the 2013 transfer supplement (which has the smallest effect, 1.6 pp) I only capture transfers above \$10,000. Standard models, including the quantiative model in this paper, predicts that renters follow a threshold rule: for any renter that chooses to rent, there exists a wealth level at which they prefer to own (Indeed, the regression results also show that a 10\% increase in net worth increases the probability of becoming a homeowner by 4 percentage points.). Hence, large transfers can shift households above the trehsold if they were `close'. Second, these transfers are not exogenous: indeed, it is likely that some of these transfers given exactly because the household is buying.

\begin{table}
	\begin{threeparttable}
		\caption{The Transition to Ownership}
		\label{tab:newowners_short}
		\small 
				\input{../tabfig/regr/newowners_short.tex}

		{\begin{footnotesize}\begin{flushleft}
		\textit{Notes:} Standard errors in parentheses. \textcolor{red}{update footnote} \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
		\end{flushleft}\end{footnotesize}}		
	\end{threeparttable}
\end{table}

To illustrate the size issue and data limitations, and the need for a model, I plot the size of transfers to renters, by whether they become homeowners in $t+2$ in Figure \ref{fig:PSID_transfer_to_renters}. The left planel only includes transfers from parents to children (from the transfer supplement) in 2013 while the right panel plots the data from the transfer/inheritance above 10k figure. There are several things to note. First, we note that, among receivers, households who transition to ownership receive larger transfers on average, in both figures. Second, the transfers in the transfer supplement is much smaller than than those in the gift/inheritance question. Indeed, the median parental transfer is just about \$1k, the mean about \$3k, and the 95th percentile is at \$10k. Third, In all measures, no matter the measure, we see that receiving renters who become owners receive much larger transfers. 
\begin{figure}
	\caption{Transfer Sizes to Renters by Transition to Ownership}\label{fig:PSID_transfer_to_renters}
	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_transfers_renters_transfers.pdf}% 
	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_transfers_renters_inheritance.pdf}%

	 {\begin{footnotesize} \textit{Notes:} The left panel uses the transfers, gifts, and inheritance variable which is available in all waves of the PSID while the right panel uses the parental transfer data only available in the 2013 transfer supplement. See main text for a discussion. Vertical dotted and dashed lines denote the median and means, respectively. The figures show the kernel density estimates by group, capped at the the 99th percentile. \end{footnotesize}}
\end{figure}

\subsubsection{Households with Wealthier Parents Are Less Likely to Be Behind on Mortgages} 
\textcolor{red}{!! [[rewrite this]] !!}
To realize the benefits of homeownership, households must be able to maintain their mortgage payments, since failure to do so can lead to distressed and forced sales. I now show that households with wealthier parents are less likely to be behind on their mortgages, even though they buy more expensive houses. The PSID has collected this data since 2009, and 3.9\% of owners aged 25 to 44 are behind, on average, while 9.2\% have been behind at least once.

The sample is limited to the first time a household is observed as owners, and the control variables are the same as in the previous regression:
\begin{equation*}
Behind_i = \beta_1 \ln(Wealth)_{p(i),t-2} + \beta_2 \ln(Inc_{i,t-2}) + \beta_3 \ln(NetWorth_{i,t-2}) + \mathbf{\gamma}{X_{i,t}} + \varepsilon_{i}.
\end{equation*}
In Column 2 of Table \ref{tab:hypo}, the outcome variable is whether households are behind on the mortgage payments the first time they are observed as owners. Parental wealth in the period before purchase decreases the probability that a household will ever be behind: A 1\% increase in parental wealth decreases the probability of being behind by 0.016 percentage points and is significant at the 1\% level. Parental wealth has a larger effect than the child's net worth. In Column 3, the outcome variable is whether first-time homeowners will ever be observed to be behind on their mortgage, and coefficients and significance levels are virtually unchanged. The sample size increases in these regressions, since households who bought before 2009 are now included.
\begin{table}
	\begin{threeparttable}
		\caption{Housing Choices and Parental Wealth}
		\label{tab:hypo}
		\small 
				\input{../tabfig/regr/hypoII.tex}

		{\begin{footnotesize}\begin{flushleft}
		\textit{Notes:} Standard errors in parentheses. `Behind' refers to whether the household is behind on a mortgage. Wealth, income, parental wealth, mortgage, family size, and house values are logged. All regressions include year and state fixed effects and control for age and age-squared of both the child and parent. Specifications 1 and 2 use OLS while specifications 3 and 4 use RE and FE, respectively. \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
		\end{flushleft}\end{footnotesize}}		
	\end{threeparttable}
\end{table}

I also report the results from two specifications where I follow households post-purchase, using random effects and household fixed effects regressions (Columns 4 and 5). Once we follow households over time, the effect of parental wealth decreases. With household fixed effects, the effect is no longer significant. However, the fixed effects regressions must be interpreted with caution, since there is little within-parent variation in wealth over time. Indeed, the random and fixed effects coefficients are almost the same (-0.008 vs -0.006). Overall, the results support the hypothesis that parental wealth decreases the probability of being behind on mortgage payments.

The probability of being behind on mortgages is decreasing in parental wealth, highlighting that parental resources appear to help new homeowners maintain ownership after purchase---in addition to relaxing borrowing constraints.

@@

The exercise follows \cite{Chetty2007} closely. The outcome of interest is the change in log housing consumption, which is set to zero for households who do not move. For households who do move, housing consumption is defined as yearly rent while renting and 5\% of the market value for owner-occupied housing \citep{Davis2008}. The sample is limited to household heads who are unemployed only once between ages 25 and 45. I divide the sample by whether a household's parents were in the top parental wealth quartile (of parents with children aged 25 to 44) at the time of unemployment and compare the two groups' housing consumption growth rates at unemployment. The results are displayed in Figure \ref{fig:housinggrowthrates}. Among households with nonwealthy parents, housing consumption decreases by about 5\% during unemployment (significant at the 95\% level). In contrast, among households with wealthy parents, housing consumption even grows slightly at unemployment. Moreover, I replicate this event study using simulated data from the quantitative model and show that the results are similar (see Section \ref{sec:externalval} and Figure \ref{fig:housinggrowthrates_model}).
\begin{figure}
	\caption{Event Study: Housing Consumption at Unemployment by Parental Wealth}\label{fig:housinggrowthrates}
	\noindent\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthpoor_both}\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthrich_both}

@@

@@

		{\begin{footnotesize}\begin{flushleft}
		\textit{Notes:} Standard errors in parentheses. \textcolor{red}{update footnote} \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
		\end{flushleft}\end{footnotesize}}		
	\end{threeparttable}
\end{table}

>>>>

To quantify the impact of parental transfers on homeownership, I then combine two distinct models: housing models, which ignore parental transfers, and models of parental transfers, which ignore housing. I build a life-cycle overlapping generations model with altruistic parents and a rent-or-own decision. Parents and children interact without commitment through both inter-vivos transfers and end-of-life bequests. In models of altruism without housing (e.g., \citealp{Altonji1997a,Barczyk2014}), the transfer motive is to increase the child's consumption when they are borrowing constrained, thus having high marginal utility of wealth. The inclusion of a frictional housing market generates two new transfer motives. First, transfers can directly alleviate credit constraints, enabling wealthier parents to help their children meet down-payment requirements. Second, future transfers provide partial insurance against future income shocks, reducing risks associated with the large illiquid investment of homeownership. Moreover, the illiquidity of housing serves as a commitment device enabling children to secure future transfers. These mechanisms differ from those examined by \citet{Barczyk2020a}, who study how elderly parents’ homeownership decisions shape nursing‑home entry, care arrangements, and intergenerational transfers. As in all transfer models, the expectation of future transfers reduces children's saving motives, potentially lowering the homeownership rate.

@@

	 {\begin{footnotesize} \textit{Notes:} The left panel uses data from the SHED with the sample restricted to current owners who report this being their first home. The question was only asked in the 2015 and 2016 waves. The right panel uses data from the AHS with the sample restricted to all owner-occupied units that did not indicate the sale of a previous home as the main source of funding. The question was discontinued after 2019. Note that the left figure has year of purchase on the horizontal axis while the right figure has survey year. Dashed lines denote 95\% confidence intervals.\end{footnotesize}}
\end{figure}

\textit{SHED:} I use SHED, an annual cross-sectional survey conducted by the Federal Reserve, to observe the share of households who funded the down payment with a loan or gift from family or friends, by year of purchase. The results are plotted in the left panel of Figure \ref{fig:motivation}. The main observation is the large increase in the role of inter-vivos transfers for homeowners since 2001. From 2001 to 2007, only 10\% to 18\% of first-time buyers received transfers, while 20\% to 40\% received transfers after 2009.

\textit{AHS:} I use the AHS, which surveys occupants of housing units, to obtain a time series on whether the down payment was mostly funded by gifts or inheritances (right panel of Figure \ref{fig:motivation}). We observe a flat pattern in the 1990s, followed by a sharp decline from 2005 to 2009, coinciding with the prevalence of mortgages featuring very low down payments (e.g., \cite{Corbae2015}), and a stark increase from 2013 to today. Minor fluctuations in the percentage of homeowners relying on gifts or inheritances reflect significant shifts among new homeowners, who comprise a small fraction of all homeowners. This is because the horizontal axis denotes the survey year, not the year of purchase.

@@

In all other waves, there is a general question of gifts and inhertitances from people outside the household. We do not know the source of these transfers (e.g., they could be from friends or an inheritance from grandparents) and hence these do not know the extent to which these captures parental assistance. Yet, they, to some extent, capture the impact of transfer receipt on household decisions, and hence I use these variables in some of the empirical results, following \cite{Lee2018}.

@@

I now utilize microdata from the PSID to examine the empirical association between transfer receipt and parental financial resources on housing outcomes. The sample is limited to households aged 25 to 44 to focus on the housing choices of young adults. I use the sum of parental net worth and income (denoted as wealth) to measure parents' financial resources in all specifications for two reasons. First, this is consistent with models of altruism, as the one used later in the paper. Second, since parents' wealth and income are highly correlated and I have few observations---fewer than 1,000 for many specifications---the standard errors are large when both are included.

\subsubsection{Transfers and the Rent-to-Own Transition}

While a growing literature documents that parental transfers increase the probability that renters become homeowners (e.g.\ \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}), I replicate this result to highlight the importance of transfers for housing outcomes. I estimate
\begin{equation}	\label{eq:transferrent}
	\Pr\left(Own_{i,t+2}=1 \mid Own_{i,t}=0\right)
	= \alpha + \beta\,Transfer_{i,t+2}
	  + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation}
where \(Own_{i,t}=0\) denotes a renter at time \(t\), \(Transfer_{i,t}=1\) if the household received a transfer between survey waves, and \(X_{i,t}\) includes log parental wealth, log household income and log household net worth, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects. The sample is limited to renting households aged 25 to 44 who have never been observed as homeowners. Table \ref{tab:newowners_short} reports the estimated coefficient on transfer ($\beta$), with the full coefficient table relegated to Appendix Table \ref{tab:newowners}. These regressions are similar to those in \cite{Lee2018}, who also use the PSID.

To measure transfer receipt, I compare two definitions. Definition 1 (narrow): Parental gifts exceeding \$500 in 2012, as reported in the PSID’s 2013 Transfer Module, limiting the sample to renters in 2011.  Definition 2 (broad): Any gift or inheritance exceeding \$10,000 reported in all waves of the PSID, or a parental gift exceeding \$10,000 in 2012. This expands the sample include all waves of the PSID, and lets us include household fixed effects, though we now only capture large transfers and include some non-parental transfers.

In a univariate regression, receiving a parental transfer is associated with a 2.9 percentage point higher probability of becoming a homeowner in 2013 among renters in 2011. With a full set of controls parental transfers are associated with a 1.2 percentage point higher probability of becoming a homeowner. This represents a substantial increase relative to the baseline rent-to-own transition rate of about 14\%, though the effect is not statistically significant.

To increase statistical power, I now use the second transfer receipt definition, which also includes non-parental gifts and bequests. This leads to a ten-fold increase in sample size, but the transfer receipt rate also decrease to around 3\%, and the median transfer size jumps from \$1,500 to \$24,200. In addition, we can now follow renters over time, enabling the inclusion of household fixed effect. In a univariate regression, transfers increase the probability of becoming a homeowner by 21.9 percentage points. Even with the full set of controls, the effect of transfer receipt on the future homeownership is large and statisticaly significant: 13.5 percentage point in the OLS regression ($p<0.001$) and a 7.9 percentage point in the FE regression  ($p<0.1$). These effects are large, similar in magnitude of the baseline rent-to-own transition rate of 18 percent. A few caveats are in order: I only capture transfers above \$10,000, and the effect of transfer receipt is likely increasing with transfer size. Second, these transfers are not exogenous: indeed, it is likely that some of these transfers are given \emph{because} households are buying a house.

\begin{table}[tb]
	\centering
	\begin{threeparttable}
		\caption{The Transition to Ownership}
		\label{tab:newowners_short}
		\small 
				\input{../tabfig/regr/newowners_short.tex}

	\end{threeparttable}
	{\begin{footnotesize}\begin{flushleft}
		\textit{Notes:} Standard errors in parentheses, clusted at the household level. \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
		\end{flushleft}\end{footnotesize}}		
\end{table}

\subsubsection{Transfer Sizes}
For additional detail on the transfer variables, I plot kernel density estimates in Figure \ref{fig:PSID_transfer_to_renters}, for renters in $t$, separated by whether they own or rent in $t+2$. The left panel only includes parental transfers exceeding \$500 from the 2013 transfer supplement. The right panel only includes transfer/inheritance exceeding \$10,000 which is included in every wave. There are two takeaways. First, households who transition to ownership receive larger transfers on average, under both transfer measures. Second, the transfers in the transfer supplement are much smaller than than those in the gift/inheritance question. Indeed, the median parental transfer is about \$1k, the mean about \$3k, and the 95th percentile is at \$10k, compared to the median gift/inheritance of \$25k, mean of \$60k, and a 95th percentile of about \$130k.

While these are the best data available to follow household over time---in particular their wealth, income, and housing choices---and transfers, the data have several some limitations. First, in the transfer supplement we only observe transfers in 2012, while the PSID is collected every other year. Hence we undercount both transfer rates and sizes. Second, in the gift/inheritance question, we do not know the source of the transfer. Third, we do not know the extent to which households report intra-family loans, nor the extent to which these are repaid.
\begin{figure}
	\caption{Transfer Sizes to Renters by Transition to Ownership}\label{fig:PSID_transfer_to_renters}
	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_transfers_renters_transfers.pdf}% 
	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_transfers_renters_inheritance.pdf}%

	 {\begin{footnotesize} \textit{Notes:} The left panel uses the transfers, gifts, and inheritance variable which is available in all waves of the PSID while the right panel uses the parental transfer data only available in the 2013 transfer supplement. See main text for a discussion. Vertical dotted and dashed lines denote the median and means, respectively. All estimates are weighted and omit the top 5\% of transfer sizes. \end{footnotesize}}
\end{figure}

\subsubsection{Households with Wealthier Parents Are More Likely to Remain Homeowners}
While the previous sections have focused on the transition into ownership, I now focus on mainting homeownership. \textcolor{red}{rewrite this better: That said, \cite{bond2021role} is an exception and shows the positive association between parental wealth and children's housing outcomes in the PSID and the Health and Retirement Study. }

First, the left panel of Figure \ref{fig:owner_hazard} plots the share of owners aged 25 to 44 in year $t$ who remain owners in period $t+\Delta$. Among first-time owners, only 80\% have retained ownership four years later. For non-first-time owners, there is still a marked decrease though it is more muted, after four years about 10\% of all owners will be renting.

To understand which variables predicts retainment of ownership, I estimate
\begin{equation}	\label{eq:regmaintain}
	\Pr\left(Own_{i,t+\Delta}=1 \mid Own_{i,t}=1\right)
	= \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t})
	  + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation}

where $Own_{i,t+\Delta}=1$ denotes an owner at time $t+\Delta$, $\beta$ and $\zeta$ measures the effect of log household net worth and parental wealth, respectively  and $X_{i,t}$ is a vector of control variables. The only changes relative to Eq. \ref{eq:transferrent} is that the outcome variable conditions on current ownership and measures future ownership, that there are no controls for transfer receipt, and that the household net worth and parental wealth are taken out of the vector of control variables. The sample is limited to households aged 25 to 44. These regressions are motivated by \cite{bond2021role}, who estimate similar equations in the PSID and the Health and Retirement Survey. Figures \ref{fig:coef_allowner} and \ref{fig:coef_ftowner} plots the two coefficients of interest, over variying lead, times and their 90\% confidence intervals for all only and first-time owners only.

Two results stand out: First, we see that parental wealth is at least as important as household net worth at all horizons, though the difference is never statistically significant. Second, we see that the impact of parental wealth is increasing over time. For example, we see that a 10\% increase in parental wealth increases the probability of retaining ownership after 8 years by 0.45 percentage points among first-time owners. These patterns will be serve as one part of model validation.
\begin{figure}[tb]
	\captionsetup[subfigure]{aboveskip=-1pt}
	\caption{Mainting Homeownership}\label{fig:homeowner_exit}
	\begin{subfigure}{0.33\textwidth}%
		\caption{Exit Out of Ownership}\label{fig:owner_hazard}%
		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_ownerexit.pdf}% 
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.33\textwidth}%
		\caption{All Owners}\label{fig:coef_allowner}%
		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_coefowner.pdf}% 
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.33\textwidth}%
		\caption{First-Time Owners}\label{fig:coef_ftowner}%
		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_coefftowner.pdf}% 
	\end{subfigure}%
	\hfill

	\caption*{\footnotesize \textit{Notes:} \textcolor{red}{write this footnote.}}
\end{figure}

\subsubsection{Households with Wealthier Parents Are Less Likely to Be Behind on Mortgages} 
% To realize the benefits of homeownership, households must be able to maintain their mortgage payments, since failure to do so can lead to distressed and forced sales. I now show that households with wealthier parents are less likely to be behind on their mortgages, even though they buy more expensive houses. The PSID has collected this data since 2009, and 3.9\% of owners aged 25 to 44 are behind, on average, while 9.2\% have been behind at least once.
One potential roadblock to maintaining ownership are mortgage delinquencies, as failure to meet mortgage payments can lead to distressed and forced sales. I now show that households with wealthier parents are less likely to be behind on their mortgages, even though they buy more expensive houses. The PSID has collected this data since 2009, and about 2\% of owners aged 25 to 44 are behind.

To understand which variables predicts whether households are behind on mortgage payments, I estimate
\begin{equation*}
\Pr(Behind_{i,t+2}) = \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t}) + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation*}
and the control variables are the same as in the previous regression. Table \ref{tab:hypo} reports the estimated coefficient on household net worth ($\beta$) and parental wealth ($\zeta$), with the full coefficient table relegated to Appendix Table \ref{tab:hypo_long}.

I first limit the sample to first-time buyers to focus on distress among new owners. In a univariate regression, a 10\% increase in parental wealth is associated with a 0.07 percentage point decrease in the probability of being behind on mortgages two years later ($p<0.1$) WWith the full set of controls, the effect increases to 0.1 percentage points, though it is no longer statistically significant. These are economically meaningful since 1.6\% of first-time owners are behind on their mortgage. Moreover, parental wealth is more predictive of future delinqucies than household net worth. In Columns 3 to 5, I instead focus on all owners (who are currently not behind), increasing the sample sizes and allowing us to include household fixed effects. In these specifications, a 10\% increase in parental wealth decreases the probability of being behind two years later by 0.09 percentage points in the univariate OLS regression ($p<0.001$), 0.02 percentage points in the OLS regression (not significant) and  by 0.07 percentage points in the FE regression ($p<0.1$). All told, the probability of being behind on mortgages is decreasing in parental wealth, highlighting that parental resources appear to help new homeowners maintain ownership after purchase---in addition to relaxing borrowing constraints.
\begin{table}
	\centering
	\begin{threeparttable}
		\caption{Housing Choices and Parental Wealth}
		\label{tab:hypo}
		\small 
				\input{../tabfig/regr/hypoII_short.tex}

	\end{threeparttable}
	{\begin{footnotesize}\begin{flushleft}
		\textit{Notes:} Standard errors in parentheses, clustered at the household level. \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
		\end{flushleft}\end{footnotesize}}	
\end{table}

@@

The exercise follows \cite{Chetty2007} closely. The outcome of interest is the change in log housing consumption, which is set to zero for households who do not move. For households who do move, housing consumption is defined as yearly rent while renting and the imputed rent of owner-occupied housing is set to 5\% of the market value \citep{Davis2008}. The sample is limited to household heads who are unemployed only once between ages 25 and 44. I divide the sample by whether a household's parents were in the top parental wealth quartile (of parents with children aged 25 to 44) at the time of unemployment and compare the two groups' housing consumption growth rates at unemployment. The results are displayed in Figure \ref{fig:housinggrowthrates}. Among households with nonwealthy parents, housing consumption decreases by about 5\% during unemployment (significant at the 95\% level). In contrast, among households with wealthy parents, housing consumption even grows slightly at unemployment. Moreover, I replicate this event study using simulated data from the quantitative model and show that the results are similar (see Section \ref{sec:externalval} and Figure \ref{fig:housinggrowthrates_model}).
\begin{figure}
	\caption{Event Study: Housing Consumption at Unemployment by Parental Wealth}\label{fig:housinggrowthrates}
	\noindent\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthpoor_both}\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthrich_both}

@@

\begin{table}
	\centering
	\begin{threeparttable}
		\caption{Housing Choices and Parental Wealth}
		\label{tab:hypo_long}
		\small 
				\input{../tabfig/regr/hypoII.tex}

	\end{threeparttable}
	{\begin{footnotesize}\begin{flushleft}
		\textit{Notes:} Standard errors in parentheses, clustered at the household level. `Behind' refers to whether the household is behind on a mortgage. Wealth, income, parental wealth, mortgage, family size, and house values are logged. All regressions include year and state fixed effects and control for age and age-squared of both the child and parent. Specifications 1 and 2 use OLS while specifications 3 and 4 use RE and FE, respectively. \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
		\end{flushleft}\end{footnotesize}}	
\end{table}

@@

		{\begin{footnotesize}\begin{flushleft}
		\textit{Notes:} Standard errors in parentheses, clustered at the household level. \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
		\end{flushleft}\end{footnotesize}}		
	\end{threeparttable}
\end{table}
